---
title: "Reproducible Research: Assignment 1"
author: "Ahmed Gomaa"
date: "19 November 2016"
output: html_document
---

## Preparation for publishing

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

## Code for reading in the dataset and/or processing the data

```{r}
data = read.csv("activity.csv")
```

## Histogram of the total number of steps taken each day

```{r}
library(ggplot2)

total.steps = tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)

qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")
```

## Mean and median number of steps taken each day

```{r}
mean(total.steps, na.rm = TRUE)

median(total.steps, na.rm = TRUE)
```

## Time series plot of the average number of steps taken

```{r}
averages = aggregate(x = list(steps = data$steps), by = list(interval = data$interval), FUN = mean, na.rm = TRUE)

ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + ylab("average number of steps taken")
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r}
averages[which.max(averages$steps), ]
```

## Code to describe and show a strategy for imputing missing data

```{r}
missing = is.na(data$steps)

table(missing)

fill.value = function(steps, interval)
             {
                filled = NA
                
                if (!is.na(steps))
                        
                        filled = c(steps)
                
                else
                        filled = averages[averages$interval == interval, "steps"]
                
                return(filled)
             }

filled.data = data

filled.data$steps = mapply(fill.value, filled.data$steps, filled.data$interval)

```

## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
total.steps = tapply(filled.data$steps, filled.data$date, FUN = sum)

qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")

mean(total.steps)

median(total.steps)
```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
weekday.or.weekend = function(date)
                     {
                        day = weekdays(date)
                        
                        if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
                        return("weekday")
                        
                        else if (day %in% c("Saturday", "Sunday"))
                                return("weekend")
                        else
                                stop("invalid date")
                     }

filled.data$date = as.Date(filled.data$date)

filled.data$day = sapply(filled.data$date, FUN = weekday.or.weekend)

```

## All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

```{r}
averages = aggregate(steps ~ interval + day, data = filled.data, mean)

ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + xlab("5-minute interval") + ylab("Number of steps")
```

